# ProPer visualization (III): *Periograms*

Adjust the periodic energy and FO curves and create their visual interaction, a.k.a. *Periogram*.

```{r clean_start}
rm(list = ls())

## Load required libraries 
require(ggplot2)
require(dplyr)
require(seewave)
require(Cairo)

raw_df <- read.csv("data_tables/raw_df.csv") %>% distinct(file, t, .keep_all = TRUE)
```

## Prepare periodic energy & F0 curves:
1. Adjust 'perFloor' and 'relTo' in first rows.
2. Use the following plots and codes to verify and change 'perFloor' values.

```{r prepare_main_df, warning=FALSE}
main_df <- raw_df
#
  #####################################
  ############## Presets ##############
  #####################################
#### set floor for log-transform (1.00 = 100%):
# adjust to find the most fitting value (change later to choose multiple values)
perFloor <- .01 # {.001 -- .05} 
#
#### anchor for relative measurements, 
# relative to the entire data ('data'), each tokeb ('token'), or speaker ('speaker'):
relTo <- c("data", "speaker", "token")[2] # {[1] -- [3]} choose 'speaker' if applicable
#
#### set floor for periodic fraction (1.00 = 100%):
# keep at 0.25 unless you need this changed
strengThresh <- 0.25 # {0 -- .5} 
#
  #####################################
  ####### compute new variables #######
  #####################################
## find maximum values for entire data set
main_df <- mutate(
  group_by(main_df),
  max_data_per_power <- max(raw_df$periodic_power, na.rm=T),
  max_data_strength <- max(raw_df$strength_rowmax, na.rm=T),
  f0_data_min <- min(raw_df$f0_smooth, na.rm=T),
  f0_data_max <- max(raw_df$f0_smooth, na.rm=T)
  )
## find maximum values for speaker-defined sets (if exist)
if(length(raw_df$speaker)>0) main_df <- mutate(group_by(main_df, speaker),
  max_speaker_per_power = max(periodic_power, na.rm=T),
  max_speaker_strength = max(strength_rowmax, na.rm=T),
  f0_speaker_min = min(f0_smooth, na.rm=T),
  f0_speaker_max = max(f0_smooth, na.rm=T),
  f0_speaker_median = round(median(f0_smooth, na.rm = T),2),
  f0_speaker_mean = round(mean(f0_smooth, na.rm = T),2),
  f0_speaker_range = round(f0_speaker_max - f0_speaker_min, 2)
  )
## find maximum values for each token 
main_df <- mutate(
  group_by(main_df, file),
  max_token_per_power = max(periodic_power, na.rm=T),
  max_token_strength = max(strength_rowmax, na.rm=T),
  f0_token_min = min(f0_smooth, na.rm=T),
  f0_token_max = max(f0_smooth, na.rm=T),
  f0_token_median = round(median(f0_smooth, na.rm = T),2),
  f0_token_mean = round(mean(f0_smooth, na.rm = T),2),
  f0_token_range = round(f0_token_max - f0_token_min, 2),
  # variables for plot normalization
  plotFloor = round(f0_token_min - f0_token_range, 2),
  ## conclude relative anchors and thresholds
  perFloorFix = perFloor, # keep records for potential fix
  perFloor_indeed = ifelse(
    relTo=="token", round(max_token_per_power * perFloor, 10), ifelse(
      relTo=="data", round(max_data_per_power * perFloor, 10), 
      round(max_speaker_per_power * perFloor, 10))),
  strengThresh_indeed = ifelse(
    relTo=="token", round(max_token_strength * strengThresh, 8), ifelse(
      relTo=="data", round(max_data_strength * strengThresh, 8), 
      round(max_speaker_strength * strengThresh, 8))),
  ## create new periodic power vector
  periodic_fraction = ifelse(strength_rowmax < strengThresh_indeed, 0, strength_rowmax),
  postPP = round(total_power * periodic_fraction, 9),
  ## log periodic power
  logPP = 10*log10(postPP/perFloor_indeed),
  logPP = ifelse(logPP<0 | is.na(logPP), 0, logPP),
  ## create relative scales (0--1) 
  intensityRel = ifelse(
    intensity<0, 0, round(intensity / max(intensity, na.rm=TRUE),7)), 
  total_powerRel = ifelse(
    total_power<0, 0, round(total_power / max(total_power, na.rm=TRUE),7)), 
  postPP_rel = ifelse(
    postPP<0, 0, round(postPP / max(postPP, na.rm=TRUE), 7)),
  logPP_rel = round(logPP / max(logPP, na.rm=TRUE),7),
  ########## periodic energy smoothing (log+smooth = smog)
  ### 20Hz low-pass filter (50ms intervals): "segmental smooth"
  smogPP_20Hz = bwfilter(wave = logPP, f = 1000, to = 20, n = 2),
  smogPP_20Hz = ifelse(
    smogPP_20Hz < 0, 0, round(smogPP_20Hz / max(smogPP_20Hz,na.rm=T), 7)),
  ### 12Hz low-pass filter (~83.3ms  intervals): "seg-syll smooth"
  smogPP_12Hz = bwfilter(wave = logPP, f = 1000, to = 12, n = 1),
  smogPP_12Hz = ifelse(
    smogPP_12Hz < 0, 0, round(smogPP_12Hz / max(smogPP_12Hz, na.rm=T), 7)),
  ### 8Hz low-pass filter (125ms  intervals): "syll-seg smooth"
  smogPP_8Hz = bwfilter(wave = logPP, f = 1000, to = 8, n = 1),
  smogPP_8Hz = ifelse(
    smogPP_8Hz < 0, 0, round(smogPP_8Hz / max(smogPP_8Hz, na.rm=T), 7)),
  ### 5Hz low-pass filter (200ms  intervals): "syllabic smooth"
  smogPP_5Hz = bwfilter(wave = logPP, f = 1000, to = 5, n = 1),
  smogPP_5Hz = ifelse(
    smogPP_5Hz < 0, 0, round(smogPP_5Hz / max(smogPP_5Hz, na.rm=T), 7)),
  ########## F0 interpolating and smooting
  f0_interp = pracma::interp1(t, f0_smooth),
  f0_interp_stretch = ifelse(
    (is.na(f0_interp) & t<min(which(!is.na(f0_interp)))), 
    f0_interp[min(which(!is.na(f0_interp)))], ifelse(
      (is.na(f0_interp) & t>=max(which(!is.na(f0_interp)))), 
      f0_interp[max(which(!is.na(f0_interp)))], f0_interp)),
  ### 6Hz low-pass filter (166.7ms  intervals): "vibrato smooth"
  f0_interp_stretch_smooth = round(bwfilter(wave = f0_interp_stretch, f = 1000, to = 6, n = 1),2),
  f0_interp_smooth = ifelse(
    !is.na(f0_interp), f0_interp_stretch_smooth, NA)
  )

```

## plot selected tokens (detailed review of the data)
Use the comment in/out (#) to toggle visualization of different data components.

```{r plot_singles, warning=FALSE, echo=FALSE}
files <- main_df$file
files <- files[!duplicated(files)==TRUE]

##################################
########### loop start ###########
plyr::ldply(files, function(f){
sel_file1 <- f
##################################

#####################################
###### manual singles, no-loop ######
# sel_file1 <- files[3] # or: "filename"
#####################################

single_token1 <- dplyr::filter(main_df, file==sel_file1)

plotUnits1 <- round(single_token1$f0_token_range[1]/30)

periogram_single1 <-
  ggplot(single_token1, aes(x=t)) +
########## F0 curves
## pre-smoothed F0 (from pitch object)
  # geom_point(aes(y=f0_row1),color="green", alpha=.5, size=.5) +
## smoothed F0 (from pitch tier)
  # geom_point(aes(y=f0_smooth),color="blue3", alpha=.3, size=.3) +
## interpolated & stretched F0 
  # geom_point(aes(y=f0_interp_stretch),color="red", alpha=.3, size=.3) +
## periogram (smogPP)
  geom_line(aes(y=f0_interp_stretch_smooth),color="magenta2", alpha=single_token1$smogPP_20Hz, size=single_token1$smogPP_20Hz*5) +
########## Power/intensity
## intensity
  # geom_line(aes(y=intensityRel*f0_token_range+plotFloor),color="yellow", alpha=.6, size=.5) +
## power
  # geom_line(aes(y=total_powerRel*f0_token_range+plotFloor),color="cornflowerblue", alpha=.5, size=.5, linetype="dashed") +
########## Periodic fraction (strength/HNR)
## raw strength (before "strengThresh")
  # geom_line(aes(y=strength_rowmax*f0_token_range+plotFloor), color="green", alpha=.2, size=.75, linetype="twodash") +
## processed strength  (after "strengThresh")
  # geom_line(aes(y=periodic_fraction*f0_token_range+plotFloor), color="tomato", alpha=.7, size=.5, linetype="dotted") +
########## Periodic power 'pp' (total power * periodic fraction)
  geom_line(aes(y=postPP_rel*f0_token_range+plotFloor),color="purple3", alpha=.5, size=.5, linetype="solid") +
########## Log periodic power 'logPP' (10*log10(PER/per_thresh))
  # geom_line(aes(y=logPP_rel*f0_token_range+plotFloor),color="seashell", alpha=.3, size=2, linetype="longdash") +
########## Smoothed logPP 'smogPP' (4 smoothing flavors: 5/ 8/ 12/ 20 Hz low-pass filter)
  geom_line(aes(y=smogPP_20Hz*f0_token_range+plotFloor),color="lightsteelblue", alpha=.5, size=.75) +
  # geom_line(aes(y=smogPP_12Hz*f0_token_range+plotFloor),color="lightyellow", alpha=.6, size=1) +
  # geom_line(aes(y=smogPP_8Hz*f0_token_range+plotFloor),color="moccasin", alpha=.4, size=1.5) +
  geom_line(aes(y=smogPP_5Hz*f0_token_range+plotFloor),color="rosybrown1", alpha=.3, size=2) +
########## TextGrids boundaries and annotations (comment out if not available)
## boundaries
  geom_vline(aes(xintercept=single_token1$syll_bounds), linetype="dotted", color="white", size=.5, alpha=.5) +
## annotations
  geom_text(aes(x=single_token1$syll_mid, y=f0_token_max+plotUnits1*2, label=as.character(syll_label), check_overlap=T), size=3, color="white", family= "Helvetica") + 
## plot stuff
  ggtitle(paste0(sel_file1)) +  
  xlab("Time (ms)") + ylab("F0 (Hz)") +
  ylim(single_token1$plotFloor[1],single_token1$f0_token_max[1]+plotUnits1*2) +
  theme(plot.title = element_text(colour = "gray"), panel.background = element_blank(), plot.background = element_rect(fill = "black"), panel.grid = element_blank(), axis.title = element_text(colour = "gray"), axis.ticks = element_blank())
print(periogram_single1)
##--save?
ggsave(periogram_single1,file=paste0("plots/",sel_file1,"_PERIOGRAM(1).pdf"),device=cairo_pdf)

##################################
############ loop end ############
})
##################################

```

## re-adjust selected tokens: change the perFloor value of specific tokens

```{r readjust_singles, warning=FALSE, echo=FALSE}

## change the perFloor of specific tokens
main_df <- mutate(
  group_by(main_df, file),
  #
  perFloorFix = ifelse(
    file == files[1],    # choose file to change
    0.004,                # set the new perFloor value
    perFloorFix),
  #                      # repeat previous 4 lines for each token if necessary
  perFloorFix = ifelse(
    file == files[3],    
    0.04,                
    perFloorFix),
  #                      
  #### re-run 
  perFloor_indeed = ifelse(
    relTo=="token", round(max_token_per_power * perFloorFix, 10), ifelse(
      relTo=="data", round(max_data_per_power * perFloorFix, 10), 
      round(max_speaker_per_power * perFloorFix, 10))),
  ## log periodic power
  logPP = 10*log10(postPP/perFloor_indeed),
  logPP = ifelse(logPP<0 | is.na(logPP), 0, logPP),
  ########## periodic energy smoothing (log+smooth = smog)
  ### 20Hz low-pass filter (50ms intervals): "segmental smooth"
  smogPP_20Hz = bwfilter(wave = logPP, f = 1000, to = 20, n = 2),
  smogPP_20Hz = ifelse(
    smogPP_20Hz < 0, 0, round(smogPP_20Hz / max(smogPP_20Hz,na.rm=T), 7)),
  ### 12Hz low-pass filter (~83.3ms  intervals): "seg-syll smooth"
  smogPP_12Hz = bwfilter(wave = logPP, f = 1000, to = 12, n = 1),
  smogPP_12Hz = ifelse(
    smogPP_12Hz < 0, 0, round(smogPP_12Hz / max(smogPP_12Hz, na.rm=T), 7)),
  ### 8Hz low-pass filter (125ms  intervals): "syll-seg smooth"
  smogPP_8Hz = bwfilter(wave = logPP, f = 1000, to = 8, n = 1),
  smogPP_8Hz = ifelse(
    smogPP_8Hz < 0, 0, round(smogPP_8Hz / max(smogPP_8Hz, na.rm=T), 7)),
  ### 5Hz low-pass filter (200ms  intervals): "syllabic smooth"
  smogPP_5Hz = bwfilter(wave = logPP, f = 1000, to = 5, n = 1),
  smogPP_5Hz = ifelse(
    smogPP_5Hz < 0, 0, round(smogPP_5Hz / max(smogPP_5Hz, na.rm=T), 7))
  )

```

## re-plot after changes (if relevant)

```{r re-plot, warning=FALSE, echo=FALSE}
files <- main_df$file
files <- files[!duplicated(files)==TRUE]

##################################
########### loop start ###########
plyr::ldply(files, function(f){
sel_file2 <- f
##################################

#####################################
###### manual singles, no-loop ######
# sel_file2 <- files[3] # or: "filename"
#####################################

single_token2 <- dplyr::filter(main_df, file==sel_file2)

plotUnits2 <- round(single_token2$f0_token_range[1]/30)

periogram_single2 <-
  ggplot(single_token2, aes(x=t)) +
########## F0 curves
## periogram (smogPP)
  geom_line(aes(y=f0_interp_stretch_smooth),color="magenta2", alpha=single_token2$smogPP_20Hz, size=single_token2$smogPP_20Hz*5) +
########## Periodic power 'pp' (total power * periodic fraction)
  geom_line(aes(y=postPP_rel*f0_token_range+plotFloor),color="purple3", alpha=.5, size=.5, linetype="solid") +
########## Log periodic power 'logPP' (10*log10(PER/per_thresh))
  # geom_line(aes(y=logPP_rel*f0_token_range+plotFloor),color="seashell", alpha=.3, size=2, linetype="longdash") +
########## Smoothed logPP 'smogPP' (4 smoothing flavors: 5/ 8/ 12/ 20 Hz low-pass filter)
  geom_line(aes(y=smogPP_20Hz*f0_token_range+plotFloor),color="lightsteelblue", alpha=.5, size=.75) +
  # geom_line(aes(y=smogPP_12Hz*f0_token_range+plotFloor),color="lightyellow", alpha=.6, size=1) +
  # geom_line(aes(y=smogPP_8Hz*f0_token_range+plotFloor),color="moccasin", alpha=.4, size=1.5) +
  geom_line(aes(y=smogPP_5Hz*f0_token_range+plotFloor),color="rosybrown1", alpha=.3, size=2) +
########## TextGrids boundaries and annotations (comment out if not available)
## boundaries
  geom_vline(aes(xintercept=single_token2$syll_bounds), linetype="dotted", color="white", size=.5, alpha=.5) +
## annotations
  geom_text(aes(x=single_token2$syll_mid, y=f0_token_max+plotUnits2*2, label=as.character(syll_label), check_overlap=T), size=3, color="white", family= "Helvetica") + 
## plot stuff
  ggtitle(paste0(sel_file2)) +  
  xlab("Time (ms)") + ylab("F0 (Hz)") +
  ylim(single_token2$plotFloor[1],single_token2$f0_token_max[1]+plotUnits2*2) +
  theme(plot.title = element_text(colour = "gray"), panel.background = element_blank(), plot.background = element_rect(fill = "black"), panel.grid = element_blank(), axis.title = element_text(colour = "gray"), axis.ticks = element_blank())
print(periogram_single2)
##--save?
ggsave(periogram_single2,file=paste0("plots/",sel_file2,"_PERIOGRAM(2).pdf"),device=cairo_pdf)

##################################
############ loop end ############
})
##################################

```

## Done?
After reviewing the data and adjusting theresholds, write the main data file for further processing

```{r minimize_main_df}

mini_main_df <- main_df[c("file","t","speaker","syll_mid","syll_bounds","syll_label","word_mid","word_bounds","word_label","f0_speaker_min","f0_speaker_max","f0_speaker_median","f0_speaker_mean","f0_speaker_range","f0_token_min","f0_token_max","f0_token_median","f0_token_mean","f0_token_range","plotFloor","intensityRel","postPP_rel","logPP_rel","smogPP_20Hz","smogPP_12Hz","smogPP_8Hz","smogPP_5Hz","f0_interp_stretch_smooth","f0_interp_smooth")]

mini_main_df <- droplevels(mini_main_df)

```

```{r write_main_df}
## Write the main data file
write.csv(mini_main_df, "data_tables/main_df.csv", row.names=FALSE)
```