# ProPer visualization (III): *Periograms*

Adjust the periodic energy and FO curves and create their visual interaction, a.k.a. *Periogram*.

```{r clean_start}
rm(list = ls())

## Load required libraries 
require(ggplot2)
require(dplyr)
require(seewave)
require(Cairo)
```

```{r read_raw_df, warning=FALSE}
main_df <- read.csv("data_tables/raw_df.csv") %>% distinct(file, t, .keep_all = TRUE)
```

## Prepare periodic energy & F0 curves

Tweak the paramenters in first 4 rows to adjust your data. Use the visual plots in next chunks to 

```{r prepare_main_df, warning=FALSE}
################
################-------- Pre-sets
################
sampleRate <- 1000                          # (100 -- 44100) number of data points per second
strengthresh <- 0.40                        # (0 -- 0.5) floor for periodic fraction (percentage from 1.00 = 100%)
per_floor <- .030                           # (.001 -- .05) floor after log-transform (percentage from 1.00 = 100%)
rel_to <- c("stim", "data", "speaker")[1]   # ([1] -- [3]) anchor for relative measurements (relative to each stimulus, the entire data set, or to pre-defined speaker sets)
#
## find maximum values for entire data set
max_data_per_power <- round(max(main_df$periodic_power, na.rm=T),7)
max_data_strength <- max(main_df$strength_rowmax, na.rm=T)
f0_data_min <- min(main_df$f0_smooth, na.rm=T)
f0_data_max <- max(main_df$f0_smooth, na.rm=T)
## find maximum values for speaker-defined sets (if exist)
if(length(main_df$speaker)>0) main_df <- mutate(group_by(main_df,speaker),
                  max_speaker_per_power = max(periodic_power, na.rm=T),
                  max_speaker_strength = max(strength_rowmax, na.rm=T),
                  f0_speaker_min = min(main_df$f0_smooth, na.rm=T),
                  f0_speaker_max = max(main_df$f0_smooth, na.rm=T))
## find maximum values for each stimulus 
main_df <- mutate(group_by(main_df,file),
                  max_stim_per_power = max(periodic_power, na.rm=T),
                  max_stim_strength = max(strength_rowmax, na.rm=T),
                  f0_stim_min = min(f0_smooth, na.rm=T),
                  f0_stim_max = max(f0_smooth, na.rm=T),
#
## conclude relative anchors and thresholds
                  per_thresh_indeed = ifelse(rel_to=="stim", max_stim_per_power * per_floor, 
                                             ifelse(rel_to=="data", max_data_per_power * per_floor, 
                                                    max_speaker_per_power * per_floor)),
                  strengthresh_indeed = ifelse(rel_to=="stim", max_stim_strength * strengthresh, 
                                               ifelse(rel_to=="data", max_data_strength * strengthresh, 
                                                      max_speaker_strength * strengthresh)),
                  ## create new periodic power vector
                  periodic_fraction = ifelse(strength_rowmax < strengthresh_indeed, 0, strength_rowmax),
                  pp_post = total_power * periodic_fraction,
                  ## log periodic power
                  log_pp = 10*log10(pp_post/per_thresh_indeed),
                  log_pp = ifelse(log_pp<0 | is.na(log_pp), 0, log_pp),
                  ## create relative scales (0--1) 
                  intRel = ifelse(intensity<0, 0, 
                                   round(intensity/max(intensity, na.rm=TRUE),7)), 
                  total_powerRel = ifelse(total_power<0, 0, 
                                           round(total_power/max(total_power, na.rm=TRUE),7)), 
                  pp_postRel = ifelse(pp_post<0, 0, 
                                                   round(pp_post / max(pp_post, na.rm=TRUE), 7)),
                  log_ppRel = round(log_pp/max(log_pp, na.rm=TRUE),7),
################
################-------- Mutations
################
                  ########## periodic energy smoothing (log+smooth = smog)
                  ### 20Hz low-pass filter (50ms intervals): "segmental smooth"
                  smog_pp20Hz = bwfilter(wave = log_pp, f = sampleRate, to = 20, n = 2),
                  smog_pp20Hz = ifelse(smog_pp20Hz < 0, 0, round(smog_pp20Hz / max(smog_pp20Hz,na.rm=T), 7)),
                  ### 12Hz low-pass filter (~83.3ms  intervals): "seg-syll smooth"
                  smog_pp12Hz = bwfilter(wave = log_pp, f = sampleRate, to = 12, n = 1),
                  smog_pp12Hz = ifelse(smog_pp12Hz < 0, 0, round(smog_pp12Hz / max(smog_pp12Hz, na.rm=T), 7)),
                  ### 8Hz low-pass filter (125ms  intervals): "syll-seg smooth"
                  smog_pp8Hz = bwfilter(wave = log_pp, f = sampleRate, to = 8, n = 1),
                  smog_pp8Hz = ifelse(smog_pp8Hz < 0, 0, round(smog_pp8Hz / max(smog_pp8Hz, na.rm=T), 7)),
                  ### 5Hz low-pass filter (200ms  intervals): "syllabic smooth"
                  smog_pp5Hz = bwfilter(wave = log_pp, f = sampleRate, to = 5, n = 1),
                  smog_pp5Hz = ifelse(smog_pp5Hz < 0, 0, round(smog_pp5Hz / max(smog_pp5Hz, na.rm=T), 7)),
                  ########## F0 interpolating and smooting
                  f0_interp = pracma::interp1(t, f0_smooth),
                  f0_interp_stretch = ifelse((is.na(f0_interp) & t<min(which(!is.na(f0_interp)))), 
                                             f0_interp[min(which(!is.na(f0_interp)))], 
                                             ifelse((is.na(f0_interp) & t>=max(which(!is.na(f0_interp)))), 
                                                    f0_interp[max(which(!is.na(f0_interp)))], f0_interp)),
                  ### 6Hz low-pass filter (166.7ms  intervals): "vibrato smooth"
                  f0_smooth_interp = ifelse(!is.na(f0_interp), 
                                            bwfilter(wave = f0_interp_stretch, f = sampleRate, to = 6, n = 1), NA),
                  ## adjust plot floor
                  floorDiff = 100 - round(min(f0_smooth_interp, na.rm=T))
                  )

```

## plot entire data sets for general overview of the f0 and periodic energy curves 

Create subsets with no more than 20 items in case the set is larger than 20.
Use the comment in/out (#) to toggle visualization of different data components.

```{r periogram_all, warning=FALSE, echo=FALSE}

periodicNRG <- main_df$smog_pp20Hz # 5Hz / 8Hz / 12Hz / 20Hz

periogram_all <- 
  ggplot(main_df, aes(x=t)) +
####### F0 curves
## smooth
  #geom_point(aes(y=f0_smooth),color="blue3", alpha=.3, size=.3) + 
## interpolated
  # geom_point(aes(y=f0_smooth_interp),color="red", alpha=.3, size=.3) +
## periogram (smog_pp)
  geom_line(aes(y=f0_smooth_interp),color="magenta2", alpha=periodicNRG, size=periodicNRG*3) +
########## Smoothed log_pp (4 smoothing flavors)
  # geom_line(aes(y=smog_pp20Hz*100-floorDiff),color="lightsteelblue", alpha=.5, size=.5) +
  # geom_line(aes(y=smog_pp12Hz*100-floorDiff),color="lightyellow", alpha=.6, size=1) +
  # geom_line(aes(y=smog_pp8Hz*100-floorDiff),color="moccasin", alpha=.4, size=1.5) +
  # geom_line(aes(y=smog_pp5Hz*100-floorDiff),color="rosybrown1", alpha=.3, size=2) +
########## Smoothed log_pp (final mutation in periogram)
  geom_line(aes(y=periodicNRG*100-floorDiff),color="slategray4", alpha=.7, size=1) +
####### TextGrids boundaries and labels (comment out if not available)
## boundaries
  geom_vline(aes(xintercept=main_df$interval_bounds), linetype="dotted", color="white", size=.5, alpha=.5) +
## annotations
  geom_text(aes(x=main_df$interval_mid, y=f0_stim_max+10, label=as.character(label), check_overlap=T), color="white", size=2, family= "Helvetica") +
## plot stuff
  facet_wrap(~file, strip.position = "top", scales = "free") +
  theme(plot.title = element_blank(), panel.background = element_blank(), plot.background = element_rect(fill = "black"), panel.grid = element_blank(), axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), strip.text = element_text(size = 8)) 
print(periogram_all)  
# ggsave(periogram_all, file="plots/periogram_all.pdf", device=cairo_pdf)

```

## plot selected stimuli (detailed review of the data)
Use the comment in/out (#) to toggle visualization of different data components.

```{r plot_singles, warning=FALSE, echo=FALSE}
files <- main_df$file
files <- files[!duplicated(files)==TRUE]
sel_file <- files[5] #"xxx" 
single_stim <- dplyr::filter(main_df, file==sel_file)

periodicNRG <- single_stim$smog_pp20Hz # 5Hz / 8Hz / 12Hz / 20Hz

periogram_single <-
  ggplot(single_stim, aes(x=t)) +
  ggtitle(paste0(sel_file)) +  
  #xlab("Time (ms)") + ylab("F0 (Hz)") +
########## F0 curves
## pre-smoothed F0 (from pitch object)
  # geom_point(aes(y=f0_row1),color="green", alpha=.5, size=.5) +
## smoothed F0 (from pitch tier)
  geom_point(aes(y=f0_smooth),color="blue3", alpha=.3, size=.3) +
## interpolated F0 +smoothed (from R)
  # geom_point(aes(y=f0_interp),color="orange", alpha=.3, size=.3) +
  # geom_point(aes(y=f0_smooth_interp),color="red", alpha=.3, size=.3) +
## periogram (smog_pp)
  geom_line(aes(y=f0_smooth_interp),color="magenta2", alpha=periodicNRG, size=periodicNRG*5) +
########## Power/intensity
## intensity
  # geom_line(aes(y=intRel*100-floorDiff),color="yellow", alpha=.6, size=.5) +
## power
  geom_line(aes(y=total_powerRel*100-floorDiff),color="cornflowerblue", alpha=.5, size=.5, linetype="dashed") +
########## Periodic fraction (strength/HNR)
## raw strength (before "strengthresh")
  # geom_line(aes(y=strength_rowmax*100-floorDiff), color="green", alpha=.2, size=.75, linetype="twodash") +
## processed strength  (after "strengthresh")
  geom_line(aes(y=periodic_fraction*100-floorDiff), color="tomato", alpha=.7, size=.5, linetype="dotted") +
########## Periodic power (total power * periodic fraction)
  # geom_line(aes(y=pp_postRel*100-floorDiff),color="purple3", alpha=.4, size=1.5, linetype="dotdash") +
########## Log periodic power (10*log10(PER/per_thresh))
  # geom_line(aes(y=log_ppRel*100-floorDiff),color="seashell", alpha=.3, size=2, linetype="longdash") +
########## Smoothed log_pp (4 smoothing flavors)
  geom_line(aes(y=smog_pp20Hz*100-floorDiff),color="lightsteelblue", alpha=.5, size=.5) +
  geom_line(aes(y=smog_pp12Hz*100-floorDiff),color="lightyellow", alpha=.6, size=1) +
  geom_line(aes(y=smog_pp8Hz*100-floorDiff),color="moccasin", alpha=.4, size=1.5) +
  geom_line(aes(y=smog_pp5Hz*100-floorDiff),color="rosybrown1", alpha=.3, size=2) +
########## Smoothed log_pp (final mutation in periogram)
  # geom_line(aes(y=periodicNRG*100-floorDiff),color="slategray4", alpha=.7, size=1) +
########## TextGrids boundaries and annotations (comment out if not available)
## boundaries
  geom_vline(aes(xintercept=single_stim$interval_bounds), linetype="dotted", color="white", size=.5, alpha=.5) +
## annotations
  geom_text(aes(x=single_stim$interval_mid, y=f0_stim_max+10, label=as.character(label), check_overlap=T), size=4, color="white", family= "Helvetica") + 
## plot stuff
  theme(plot.title = element_text(colour = "gray"), panel.background = element_blank(), plot.background = element_rect(fill = "black"), panel.grid = element_blank(), axis.title = element_blank(), axis.text.x = element_text(), axis.ticks = element_blank()) 
print(periogram_single)
# ggsave(periogram_single,file=paste0("plots/",sel_file,"_plot.pdf"),device=cairo_pdf)

```

## Done?
After reviewing the data, adjusting all theresholds and correcting all the F0 curves, write the main data file for further processing

```{r write_main_df}
## Write the main data file
write.csv(main_df, "data_tables/main_df.csv", row.names=FALSE)
```